---
title: "GLM"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_caption: yes
    theme: cerulean
    self_contained: true
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
library(foreign)
library(arm)
library(data.table)
library(car)
library(MASS)
library(faraway)
library(olsrr) 
library(leaps)
library(DTRreg)
library(ggplot2)
library(dplyr)
library(tidyr) 
library(tidyverse)
library(ggplot2)
library(ggcorrplot)
```

In this biological project, we are interested in analyzing the effects of different characteristics of female crabs on the number of their satellites. Satellites are additional male crabs that reside near a female crab whose nest is shared. We considered various characteristics such as color, spine structure, width, weight, and the presence of satellites. The dataset contains both continuous and categorical variables.

```{r, include=FALSE}
path <-"/Users/elhambahrampour/Desktop/Portfolio Projects/Project 1 - Data Analysis - GLM"
setwd(path)
data <- read.csv("crabs2.csv")

satell<- data$satell
color <- as.factor(data$color)
spine <- as.factor(data$color)
width <- data$width
weight <- data$weight
data$sat <- ifelse(data$satell==0, 0, 1); sat <- data$sat
data$satName <- ifelse(data$satell==0, "No", "Yes"); SatName <- data$satNAmeame
```

## Correlation

We begin by creating a correlation plot to explore the relationships between these variables. As shown in the following plot, a strong relationship exists between weight and width. The plot also suggests a  relationship between the color and spine structure of the crabs.

```{r, echo=FALSE}
var <- cbind(color, spine, width, weight, sat)
corr <- cor(var)
#round(corr, 2)
#ggcorrplot(corr, hc.order = TRUE, type = "lower")
p.mat <- cor_pmat(var)
ggcorrplot(corr, hc.order = TRUE, type = "lower", p.mat = p.mat)
```

To investigate the possibility of an outlier, we examine the relationship between weight and width by plotting these two variables for each satellite, as shown below:

```{r, echo=FALSE}
gg <- ggplot(data) + 
  geom_point(aes(x=width, y=weight, color=satName)) +
  labs(x="Width", y="Weight") +
  scale_color_discrete(name="Satellite"); gg
```

## Model selection

No evidence of an outlier is present, and given the correlation between width and weight, including both variables in the model would not result in a significant contribution from the weight variable. We proceed by fitting a logistic regression model to the data. To select the appropriate model, we employ the AIC and BIC criteria using the backward, forward, and stepwise methods.

```{r, eval=FALSE}
mod <- glm(sat ~ as.factor(color) + as.factor(spine) + width
          + as.factor(color) * as.factor(spine)
          + as.factor(color) * width + as.factor(spine) * width 
          + as.factor(color) * as.factor(spine) * width, 
          family = binomial, data = data)

step(mod)      #BACKWARDS-AIC 

backward <- step(mod , k = log(13) )  #BACKWARDS-BIC

mod2 <- glm(sat ~ 1, family = binomial, data = data)
forward <- step(mod2, scope = list ( lower = y ~ 1, 
                upper = formula (mod)), direction="forward") #FORWARD

stepwise<-step(mod2, scope = list ( lower = y ~ 1, 
               upper = formula (mod)), direction="both", k = log(13) ) #STEPWISE
```
Based on the results, we ended up with two following models:

$$model1: sat \sim width$$
$$model1: sat \sim width+color$$

 
## Prediction model
Now, we compare these two models by ANOVA and can conclude that at the $\alpha-$level $0.01$ the model $sat \sim width$ is the appropriate model with the $AIC=192.66.$ The Normal QQ plot do not also reveal any problem. We can now estimate the coefficients and use this model as a prediction equation.

```{r}
mod1 <- glm(sat ~  width, family = binomial, data = data)
mod2 <- glm(sat ~  width + as.factor(color), family = binomial, data = data)
anova(mod1, mod2, test = "Chisq")
qqnorm(residuals(mod1), main="Normal Plot of Residuals")
summary(mod1)
```

Thus, we predict the probability $\pi$ of obtaining a satellite by following equation $$\mathrm{log(\frac{\pi}{1-\pi})}=-12.334+0.497\texttt{width}.$$

The following plot demonstrates the statistical model to calculate the probability of obtaining a satellite. This model uses the width of a crab. We now utilize the model to calculate the likelihood of obtaining the satellite for each crab. 

```{r}
width <- seq(20, 35, by=0.1)
prob <- exp(-12.334+0.497*width)/(1+exp(-12.334+0.497*width))
pred <- data.frame(cbind(width, prob))
ggplot(pred) + 
  geom_point(aes(x=width, y=prob)) +
  labs(title="Prediction Model", x="Width (cm)", y="Probability of having at least one satellite") + 
  theme(plot.title=element_text(size=15, face="bold"), 
                  axis.text.x=element_text(size=10), 
                  axis.text.y=element_text(size=10),
                  axis.title.x=element_text(size=12),
                  axis.title.y=element_text(size=12)) 
```

